---
title: "About"
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

Describe your project.
---# Merfish Data Analysis

Discover an advanced approach to analyzing Merfish data, a revolutionary molecular imaging technique that maps RNA in cells with unparalleled spatial precision.

## Loading and Cleaning Data

Using specialized libraries to load and clean Merfish data, removing irrelevant information.

```{r}
# Load necessary libraries
library(Seurat)
library(KODAMAextra)
library(ggplot2)
library(NMI)
library(mclust)
library(bluster)
library(igraph)

# Set working directory
setwd("~/Merfish giotto/Data/merfish")

# Load data
data <- read.csv("C:/Users/T0087231/Videos/EBTESSAM/Moffitt_and_Bambah-Mukku_et_al_merfish_all_cells.csv")
data <- as.data.frame(data)

# Data cleaning
colnames_data <- colnames(data)
blankgenes <- grep("Blank", colnames_data)
fos_index <- which(colnames(data) == "Fos")
df <- data[, -c(fos_index, blankgenes)]
selected_bregma <- unique(df$Bregma)

# Filter data for Animal_ID 1 and specific Bregma values
exp <- subset(df, Animal_ID == 1 & Bregma %in% selected_bregma)
rownames(exp) <- exp$Cell_ID
data_list <- split(exp, exp$Bregma)
data_list <- data_list[match(selected_bregma, names(data_list))]
```

## Normalization and Dimensionality Reduction

Normalization and dimensionality reduction using Principal Component Analysis (PCA) for efficient visualization.

```{r}
# Normalization and dimensionality reduction by PCA
xy <- list()
cons <- list()
for (i in selected_bregma) {
  x <- data_list[[i]]$Centroid_X - min(data_list[[i]]$Centroid_X)
  y <- data_list[[i]]$Centroid_Y - min(data_list[[i]]$Centroid_Y)
  xy[[i]] <- cbind(x, y)
  rownames(xy[[i]]) <- rownames(data_list[[i]])
  
  cons[[i]] <- t(data_list[[i]][, 10:ncol(data_list[[i]])])
  colnames(cons[[i]]) <- rownames(data_list[[i]])
  
  v <- t(LogNormalize(cons[[i]]))
  pca <- prcomp(v, scale. = TRUE)$x[, 1:50]
}
```

## KODAMA Clustering and Graph-based Clustering

Grouping data into clusters using the KODAMA algorithm, followed by refinement with graph-based clustering for accurate classification.

```{r}
# KODAMA clustering
vis <- list()
pred <- list()
refine <- list()
ARI <- list()
NMI <- list()
for (i in selected_bregma) {
  kk <- KODAMA.matrix.parallel(pca, spatial = xy[[i]], FUN = "PLS", landmarks = 10000, n.cores = 4)
  vis[[i]] <- KODAMA.visualization(kk, method = "UMAP")
  names(vis[[i]]) <- names(data_list[[i]])
  
  g <- bluster::makeSNNGraph(as.matrix(vis[[i]]), k = 100)
  g_walk <- igraph::cluster_walktrap(g)
  pred[[i]] <- as.character(igraph::cut_at(g_walk, no = 20))
  refine[[i]] <- refinecluster(pred[[i]], xy[[i]], shape = "hexagon")
  pred[[i]] <- as.character(igraph::cut_at(g_walk, no = 8))
  refine[[i]] <- refinecluster(pred[[i]], xy[[i]], shape = "hexagon")
  
  # Calculate ARI and NMI
  ARI[[i]] <- adjustedRandIndex(zlabels, refine[[i]])
  NMI[[i]] <- NMI(zlabels, refine[[i]])
}
```

## Cluster Visualization

Visualizing clusters with UMAP, a dimensionality reduction method for better data understanding.

```{r}
# Visualize clusters
plotClustersFacet(xy, pred, selected_bregma, size = 0.2) +
  scale_color_manual("Domain", values = cols) +
  guides(color = guide_legend(nrow = 1, override.aes = list(size = 2)))

plotClustersFacet(xy, refine, selected_bregma, size = 1) +
  scale_color_manual("Domain", values = cols) +
  guides(color = guide_legend(nrow = 1, override.aes = list(size = 2)))
```

## Cell Types

Identifying and visualizing different cell types for in-depth analysis.

```{r}
# Cell types visualization
plotClustersFacet(xy, clabels, selected_bregma, size = 0.2) +
  scale_color_manual("Type", values = cols) +
  guides(color = guide_legend(byrow = T, nrow = 2, override.aes = list(size = 2)))

plotCellTypes(xy[[3]], pred[[3]], cols, ncol = 5, dotsize = 0.5,
      size = 0.6, alpha = 0.5, stroke = 0) +
  ggtitle(names(data_list)[3])

plotCellTypes(xy[[3]], refine[[3]], cols, ncol = 5, dotsize = 0.5,
              size = 0.6, alpha = 0.5, stroke = 0) +
  ggtitle(names(data_list)[3])
```

This script illustrates a standard approach to Merfish data analysis, showcasing the use of commonly used bioinformatics tools. It provides a comprehensive overview of the Merfish data analysis and visualization process, crucial for understanding the spatial distribution of RNA in cells.
```
